The ChEMBL database as Linked Open Data

### 2012-04-02 article.tex
Restructured things

To expose that ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
hosted at Uppsala University, using the Open Source Virtuoso software. Use is free, but the
querying is capped, based on the estimated computational effort. Second, resources have
been made dereferencable using the Kasabi platform. This platform also provides a SPARQL
end point, but requires a user account.

@@

\subsection{Data Statistics}

@@

The ChemSpider co-authors complemented the original ChEMBL data with ChemSpider identifiers
for the chemical structures in the data set. Links were generated by ....

Valery: FIXME

Resulting in 1073967 ChemSpider identifiers.

\subsection{Data Structure}

@@

@@

\subsubsection{Bio2RDF}

@@

\subsubsection{CitedIn}

@@

\subsubsection{Decision Support}

@@

\subsection{Linked Data on Kasabi}

Later, to satisfy the requirements for the Linked Open Drug Data \cite{Samwald2011}
network, the data was also hosted on Kasabi, a triple store hosting service
free for Open Data~\cite{kasabi}.

David: Chem2Bio2RDF

Antony: if you can contribute ChemSpider IDs for the ChEMBL compounds ... 
AJW> CC-BY-SA 3.0 licensing approved for the mappings between ChEMBL and ChemSpider. Please use the Open PHACTS data provided and attribution to ChEMBL, OPS and ChemSPider.

RSC’s ChemSpider~\cite{Pence2010} is a free online database of over 26 million unique
chemical compounds aggregated from over 400 data sources as well as chemical data extracted
from RSC scientific articles and databases. Since its inception efforts have been made to
utilize the platform as both a deposition platform for the community to contribute data as
well as a platform for annotation and curation. Studies have shown that there are data
quality issues in many of the public compound databases~\cite{Williams2011} and ChemSpider has become a
valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
is presently providing the chemical services underpinning the Open PHACTS semantic web
project~\cite{Williams2012} providing access to structure, substructure and similarity searching services
to the core architecture of the project. Specific chemical data sources containing data
mappings between ChemSpider identifiers (CSIDs) and the original data source identifiers
have been provided to the triple store, together with chemical identifiers including
validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
The data mappings between the ChemSpider IDs and the data source IDs are released to
the community under Creative Commons licenses (CC-BY-SA 3.0). Link mappings of the
form:

\begin{tiny}
\begin{verbatim}
<http://rdf.chemspider.com/370>
 skos:exactMatch
 <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846>
\end{verbatim}
\end{tiny}

have been constructed.

\subsubsection{Linking out to ChEMBL}
Peter: write up the proxying... if that still works...

John/Anna: I can easily add links to webpages using foaf:homepage... are webpage URLs for entities stable enough?

@@

>>>>

To expose that ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
hosted at Uppsala University, using the Open Source Virtuoso software. Use is free, but the
querying is capped, based on the estimated computational effort. Second, resources have
been made dereferencable using the Kasabi platform~\cite{kasabi}. This platform also provides a SPARQL
end point, but requires a user account.

@@

We here present the updated ChEMBL-RDF.

\subsection{ChEMBL-RDF}

\subsubsection{Data Statistics}

@@

\subsubsection{Added Data}

RSC's ChemSpider~\cite{Pence2010} is a free online database of over 26 million unique
chemical compounds aggregated from over 400 data sources as well as chemical data extracted
from RSC scientific articles and databases. Since its inception efforts have been made to
utilize the platform as both a deposition platform for the community to contribute data as
well as a platform for annotation and curation. Studies have shown that there are data
quality issues in many of the public compound databases~\cite{Williams2011} and ChemSpider has become a
valuable resource for curated data, especially chemical-compound name mappings. ChemSpider
is presently providing the chemical services underpinning the Open PHACTS semantic web
project~\cite{Williams2012} providing access to structure, substructure and similarity searching services
to the core architecture of the project. Specific chemical data sources containing data
mappings between ChemSpider identifiers (CSIDs) and the original data source identifiers
have been provided to the triple store, together with chemical identifiers including
validated chemical names (systematic, generic and trivial), SMILES () and InChIs (). 
The data mappings between the ChemSpider IDs and the data source IDs are released to
the community under Creative Commons licenses (CC-BY-SA 3.0; attribution should be
made to the original ChEMBL database, Open PHACTS, and ChemSpider).

Link mappings are provided with skos:exactMatch predicates, while the ChemSpider identifiers
are also available via a CHEMINF representation:

\begin{tiny}
\begin{verbatim}
<http://rdf.chemspider.com/370>
 skos:exactMatch
 <http://data.kasabi.com/dataset/chembl-rdf/chemblid/CHEMBL324846> .
\end{verbatim}
\end{tiny}

FIXME: add ChemSpider IDs with CHEMINF.

\subsubsection{Data Structure}

@@

\subsubsection{Linking out to RDF OpenMolecules}

\subsubsection{Linking out to CrossRef}

@@

\section{Applications}

\subsection{Bio2RDF}

@@

\subsection{CitedIn}

@@

\subsection{Decision Support}

@@

\subsection{Chem2Bio2RDF}

David/Bin: Chem2Bio2RDF

@@

FIXME: add IMI
